---
title: "Import and format LTER data"
author: "CRF"
date: "2023-01-24"
output: html_document
editor_options: 
  chunk_output_type: console
---

#**LTER data downloaded January 23, 2023 from mcr-lter website (http://mcrlter.msi.ucsb.edu/cgi-bin/showDataset.cgi?docid=knb-lter-mcr.8&displaymodule=entity&entitytype=dataTable&entityindex=1)
#Jan 29, 2024 read a notice that the 2020 data is unreliable

The goal of this script is to import the lter data and make it useable for understanding macroalgal community dynamics

```{r pivot into long format and filter for just the algae}
library(tidyverse)

#read in data
benthic_data<-read.csv("~/Dropbox/bbj/MCR_LTER_Annual_Survey_Benthic_Cover_20220311.csv")
#remove redundant columns
benthic_data<-benthic_data[,c(1, 4, 5, 6, 7, 8, 9)]
#label columns
colnames(benthic_data)<-c("year", "site", "habitat", "trans", "quad", "taxa", "per_cover")
#remove lines where no data was entered (n=1)
benthic_data<-benthic_data%>%drop_na(per_cover)
#remove 2020 data
benthic_data<-subset(benthic_data, year!=2020)
unique(benthic_data$year) #check
#look at the taxa
unique(benthic_data$taxa)
#make the data long format
benthic_data<-
  benthic_data%>%
  pivot_wider(names_from=taxa, 
              values_from=per_cover)
#fill empty spots with 0 coer
benthic_data[is.na(benthic_data)] <- 0
#relabel the column names
algae<-benthic_data[, c("year", "site", "habitat", "trans", "quad",
                        "Amansia rhodantha", "Turbinaria ornata",  
                        "Dictyota sp.","Halimeda sp.", "Asparagopsis taxiformis",
                        "Lobophora variegata", "Peyssonnelia inamoena", "Valonia aegagropila",
                        "Dichotomaria marginata","Dichotomaria obtusata", 
                        "Amphiroa fragilissima", "Dictyota friabilis", "Galaxaura sp.",
                        "Caulerpa serrulata", "Liagora ceranoides", "Halimeda minima",
                        "Dictyota bartayresiana", "Galaxaura rugosa","Halimeda opuntia",
                        "Halimeda discoidea","Caulerpa racemosa","Galaxaura filamentosa",
                        "Valonia ventricosa", "Cladophoropsis membranacea","Martensia elegans",
                        "Halimeda distorta", "Dictyosphaeria versluysii",  
                        "Actinotrichia fragilis","Microdictyon umbilicatum", 
                        "Halimeda incrassata", "Jania sp.", "Dictyosphaeria cavernosa", 
                        "Halimeda macroloba", "Dictyota hamifera","Gelidiella acerosa",
                        "Caulerpa peltata" , "Dictyota implexa", "Peyssonnelia bornetii",
                        "Microdictyon okamurae","Halimeda taenicola", 
                        "Chlorodesmis fastigiata","Phyllodictyon anastomosans",
                        "Chnoospora implexa","Codium geppiorum", "Gibsmithia hawaiiensis",
                        "Peyssonnelia sp.","Cladophoropsis luxurians",
                        "Neogoniolithon frutescens","Neomeris vanbosseae",
                        "Caulerpa pickeringii","Rhipidosiphon javensis","Sargassum pacificum",
                        "Padina boryana", "Hypnea spinella","Chaetomorpha antennina",
                        "Boodlea kaeneana","Acanthophora spicifera", "Coelothrix irregularis",
                        "Gelidiella sp.","Hydroclathrus clathratus", "Dictyota divaricata" ,
                        "Colpomenia sinuosa")]
#remove anly lines with no algae
algae[is.na(algae)] <- 0

```
62 species of macroalgae. It is likely a good idea to filter out the ones that are super rare.
1) filter to back and fringing reefs
2) calculate percent of macroalgal community
4) check for commonality to develop some threshold

currently at 62 species.
```{r }
library(dplyr)
#sum the algae percent
algae$total <- rowSums( algae[,6:67] )
#remove all but backreef
algae<-subset(algae, habitat!=c("Outer 10")) 
algae<-subset(algae, habitat!=c("Outer 17")) 
#algae<-subset(algae, habitat!=c("Fringing")) #make sure this is what you want!

#make sure all rows with no data are gone
algae<-subset(algae, total!=0) 

#461 data points

#calculate the percent of the benthos that is that algae
per<-algae%>%mutate_at(vars(6:68),list(percent=~./total*100))
#extract just the columns with % cover
per<-per[69:130]
#rename the columns
colnames(per)<-c("Amansia rhodantha", "Turbinaria ornata",  
                        "Dictyota sp.","Halimeda sp.", "Asparagopsis taxiformis",
                        "Lobophora variegata", "Peyssonnelia inamoena", "Valonia aegagropila",
                        "Dichotomaria marginata","Dichotomaria obtusata", 
                        "Amphiroa fragilissima", "Dictyota friabilis", "Galaxaura sp.",
                        "Caulerpa serrulata", "Liagora ceranoides", "Halimeda minima",
                        "Dictyota bartayresiana", "Galaxaura rugosa","Halimeda opuntia",
                        "Halimeda discoidea","Caulerpa racemosa","Galaxaura filamentosa",
                        "Valonia ventricosa", "Cladophoropsis membranacea","Martensia elegans",
                        "Halimeda distorta", "Dictyosphaeria versluysii",  
                        "Actinotrichia fragilis","Microdictyon umbilicatum", 
                        "Halimeda incrassata", "Jania sp.", "Dictyosphaeria cavernosa", 
                        "Halimeda macroloba", "Dictyota hamifera","Gelidiella acerosa",
                        "Caulerpa peltata" , "Dictyota implexa", "Peyssonnelia bornetii",
                        "Microdictyon okamurae","Halimeda taenicola", 
                        "Chlorodesmis fastigiata","Phyllodictyon anastomosans",
                        "Chnoospora implexa","Codium geppiorum", "Gibsmithia hawaiiensis",
                        "Peyssonnelia sp.","Cladophoropsis luxurians",
                        "Neogoniolithon frutescens","Neomeris vanbosseae",
                        "Caulerpa pickeringii","Rhipidosiphon javensis","Sargassum pacificum",
                        "Padina boryana", "Hypnea spinella","Chaetomorpha antennina",
                        "Boodlea kaeneana","Acanthophora spicifera", "Coelothrix irregularis",
                        "Gelidiella sp.","Hydroclathrus clathratus", "Dictyota divaricata" ,
                        "Colpomenia sinuosa")

#reassemble the dataset
algae_data<-cbind(algae[1:5], per)


```


```{r aggregate by quadrant and then check for abundances}
library(dplyr)

#reformat data to long form
algae_long<-
  algae_data%>%
  pivot_longer(cols=6:67, 
               names_to="taxa",
              values_to="per_cover") #289863

#calculate averages based on transect (dropping quadrant)
algae_long<-algae_long %>% group_by(year,site, habitat, trans, taxa) %>% 
  summarise(cover=mean(per_cover)) %>%
  as.data.frame() #53444

#now we are going to look at % of TOTAL benthos (not just % of algal community) to generate thresholds/cuttoffs

#pivot algae dataset
thresholds<-
  algae%>%
  pivot_longer(cols=6:67, 
               names_to="taxa",
              values_to="per_cover") #150660

#calcualte averages based on transect (droping quadrant)
thresholds<-thresholds %>% group_by(year,site, habitat, trans, taxa) %>% 
  summarise(cover=mean(per_cover)) %>%
  as.data.frame() #27962

#max levels
max<-aggregate(thresholds$cover, by = list(thresholds$taxa), max)
colnames(max)<-c("species", "max_cover")
max<-max[order(max$max_cover,decreasing=FALSE),]
max
#setting to a minimum of 5% transect
max<-subset(max, max_cover>=5)
#list of species to keep
species_to_keep<-c(max$species)
##by dropping 2020 data, we lose padina!
#back to the dataset calculated as % of algal community
algae_long<-subset(algae_long, taxa %in% species_to_keep)

unique(algae_long$taxa)

algae_wide<-
  algae_long%>%
  pivot_wider(names_from=taxa, 
              values_from=cover)

```

#reduction to 28 species.Some details: 
--> combine the two peysionellas
--> need to decide what to do with Dictyota sp and Halimeda sp. Option 1: take means of their congeners (D. bartayresiana is the only one but we also have friabilis... H. discoidea, distorta, minima, incrassata, opuntia but we also have macroloba, taenicola, tuna). Option 2: make all  dictyota sp and halimeda sp
--> h. distorta is a synonym for h. incressata

#combine peyssonnelia
#dictyota (putatively barya)
#genera mean for halimeda sp, sensitivity tests to confirm

```{r reducing repeated genera}

algae_wide$`Dictyota bartayresiana`<-algae_wide$`Dictyota bartayresiana`+algae_wide$`Dictyota sp.`
algae_wide<-(algae_wide[,!names(algae_wide) %in%  c("Dictyota sp.")])

algae_wide$`Peyssonnelia spp`<-algae_wide$`Peyssonnelia bornetii`+algae_wide$`Peyssonnelia inamoena`+algae_wide$`Peyssonnelia sp.`
algae_wide<-(algae_wide[,!names(algae_wide) %in%  
                          c("Peyssonnelia bornetii","Peyssonnelia inamoena","Peyssonnelia sp.")])

algae_wide$`Halimeda incrassata`<-algae_wide$`Halimeda incrassata`+algae_wide$`Halimeda distorta`
algae_wide<-(algae_wide[,!names(algae_wide) %in%  c("Halimeda distorta")])

algae_sp<-algae_wide[5:27] #28 if include fringing

new_order = sort(colnames(algae_sp),decreasing=FALSE)
df2 <- algae_sp[, new_order]

algae_wide<-cbind(algae_wide[1:4], df2)

mcr_data<-algae_wide
````