---
title: "5_visualization"
author: "CRF"
date: "2023-02-02"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r Backreef}
library(viridis)
library(ggpubr)
library(vegan)
library(ggdensity)

en = envfit(NMDS[,5:6], year_by_traits[,5:9], permutations = 999, na.rm = TRUE)
en_coord_cont = as.data.frame(scores(en, "vectors")) #* ordiArrowMul(en)

Trait<-c("H", "DW:WW", "T", "SA:DW", "H:WW")
loading<-(cbind(Trait, NMDS1=en_coord_cont$NMDS1, NMDS2=en_coord_cont$NMDS2))
loading<-as.data.frame(loading)
loading$NMDS1<-as.numeric(loading$NMDS1)
loading$NMDS2<-as.numeric(loading$NMDS2)


colfunc <- colorRampPalette(c("#C0C0C0", "#045D5D"))
colors<-colfunc(10)


egg<-ggplot()+
    xlim(-1, 1)+ylim(-1, 1)+
    scale_fill_continuous(low = "#578989", high = "#F6BE00")+
stat_density_2d(data=subset(NMDS, habitat=="Backreef"), geom = "polygon", aes(x=NMDS1, y=NMDS2, fill = after_stat(level)), bins=100,  show.legend=FALSE, alpha=1)+
  geom_text(data=loading, aes(x=NMDS1, y=NMDS2, label=Trait),hjust=0, vjust=0)+
  geom_hline(yintercept=0, linetype="dashed", linewidth=0.25, color="grey")+ 
  geom_vline(xintercept=0, linetype="dashed", linewidth=0.25, color="grey")+ 
  geom_segment(data=loading, aes(xend=NMDS1, yend = NMDS2), linewidth=.5, arrow = arrow(length = unit(0.25, "cm")),
               x = 0, y = 0, color = "black") +
  geom_point(data=NMDS, aes(x=NMDS1, y=NMDS2), alpha=0.25, size=.25)+
  theme_classic()

xmax<-max(cent_2$NMDS1)
ymax<-max(cent_2$NMDS2)
xmin<-min(cent_2$NMDS1)
ymin<-min(cent_2$NMDS2)

#1200x500

colfunc <- colorRampPalette(c("#F6BE00", "#045D5D"))

colors<-colfunc(17)

backreef_all<-ggplot(data=subset(cent_2, habitat=="Backreef"),
                 aes(x=NMDS1, y=NMDS2, color=year, group=habitat))+
  geom_point(size=2)+
  geom_path()+
  scale_color_manual(values=colors)+
  xlim(xmin, xmax)+ylim(ymin, ymax)+
  geom_hline(yintercept=0, linetype="dashed", linewidth=0.25)+ 
  geom_vline(xintercept=0, linetype="dashed", size=0.25)+ 
  geom_text(aes(label=year),hjust=0, vjust=0)+
  theme_classic()+theme(legend.position = "none")

fringing_all<-ggplot(data=subset(cent_2, habitat=="Fringing"),
                 aes(x=NMDS1, y=NMDS2, color=year, group=habitat))+
  geom_point(size=2)+
  geom_path()+
  scale_color_manual(values=colors)+
  xlim(xmin, xmax)+ylim(ymin, ymax)+
  geom_hline(yintercept=0, linetype="dashed", linewidth=0.25)+ 
  geom_vline(xintercept=0, linetype="dashed", size=0.25)+ 
  geom_text(aes(label=year),hjust=0, vjust=0)+
  theme_classic()+theme(legend.position = "none")

ggarrange(loading_plot, backreef_all, fringing_all, ncol=1)


```

```{r}
library(corrplot)

correlation<-cor(year_by_traits[5:9])
cor<-corrplot(correlation, method="color", type="upper",order="hclust", tl.col="black", addCoef.col = 'black', number.cex = 0.8)#1000
cor(year_by_traits[5:10])

```

LOESS smoothing is a non-parametric form of regression that uses a weighted, sliding-window, average to calculate a line of best fit. Within each "window", a weighted average is calculated, and the sliding window passes along the x-axis.

One can control the size of this window with the span argument. The span element controls the alpha, which is the degree of smoothing. The smaller the span, the smaller the 'window', hence the noisier/ more jagged the line.

Look for documentation under LOESS rather than span.

```{r}

normalized = (x-min(x))/(max(x)-min(x))

df<-cbind(NMDS1=NMDS$NMDS1, NMDS2=NMDS$NMDS2,
          H=((year_by_traits$H-min(year_by_traits$H))/(max(year_by_traits$H)-min(year_by_traits$H))), 
          DW.WW=((year_by_traits$DW.WW-min(year_by_traits$DW.WW))/(max(year_by_traits$DW.WW)-min(year_by_traits$DW.WW))), 
          P=((year_by_traits$P-min(year_by_traits$P))/(max(year_by_traits$P)-min(year_by_traits$P))),
          SA.DW=((year_by_traits$SA.DW-min(year_by_traits$SA.DW))/(max(year_by_traits$SA.DW)-min(year_by_traits$SA.DW))),
          H.WW=((year_by_traits$H.WW-min(year_by_traits$H.WW))/(max(year_by_traits$H.WW)-min(year_by_traits$H.WW))))

df<-as.data.frame(df)

min(df$NMDS1)
NMDS1_reg<-ggplot(df)+
  geom_smooth(aes(NMDS1, H, color="Height"))+
  geom_smooth(aes(NMDS1, DW.WW, color="DW:WW"))+
  geom_smooth(aes(NMDS1, P, color="P"))+
  geom_smooth(aes(NMDS1, SA.DW, color="SA:DW"))+
  geom_smooth(aes(NMDS1, H.WW, color="H:WW"))+
  ylab("Value")+ xlim(-1, 1)+
  theme_classic() + theme(legend.position="bottom")



NMDS2_reg<-ggplot(df)+
  stat_smooth(aes(NMDS2, H, color="Height"))+
  geom_smooth(aes(NMDS2, DW.WW, color="DW:WW"))+
  geom_smooth(aes(NMDS2, P, color="P"))+
  geom_smooth(aes(NMDS2, SA.DW, color="SA:DW"))+
  geom_smooth(aes(NMDS2, H.WW, color="H:WW"))+
  ylab("Value")+xlim(-1, 1)+ylim(0, 1)+
  theme_classic()+ theme(legend.position="bottom")

loading_plot #6 x 5
NMDS1_reg
NMDS2_reg


```
